---
title: "Unit 9 Assignment"
author: "Steve Bramhall"
date: "October 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # each code chunk will have their own echo control
```

```{r message=FALSE, warning=FALSE}
# Load Libraries
library(dplyr)      # for string functions
library(ggplot2)    # for plots
library(kableExtra) # for table formatting
library(knitr)      # for presenting in html
library(stringr)    # for str_trim function

```

###A. Clean and prepare the data

####1. First read the two data sets into data frames beerDF and breweryDF from csv files. The columns (variables) of each data frame were custom named to be more readable and to provide a handle to merge the data.Create column for Brewery ID that is common to both datasets.
```{r echo = TRUE}
# Read beer data from csv file and put into breweryDF data frame.Then rename columns to be more readable.
beerDF <- read.csv(".\\InputFiles\\Beers.csv",header=TRUE,sep=",",stringsAsFactors = TRUE, encoding = "UTF-8")
names(beerDF) <- c("BeerName","BeerID","AlcoholByVolume","BitternessUnits","BreweryID",
                   "BeerStyle","Ounces") # rename table columns

# Read brewery data from csv file and put into breweryDF data frame.Then rename columns to be more readable.
breweryDF <- read.csv(".\\InputFiles\\Breweries.csv",header=TRUE,sep=",",stringsAsFactors = FALSE)
names(breweryDF) <- c("BreweryID","BreweryName","City","State") # rename table columns
```

####2. Merge the beer and brewery data sets into a sigle dataframe.
```{r echo = TRUE}
#Merge brewery and beer data into the brewDataDF data frame
allBeerDataDF <- merge.data.frame(breweryDF,beerDF,by="BreweryID")
```

####3. Remove surrounding whitespace for State info
```{r echo = TRUE}
allBeerDataDF$State <- sapply(allBeerDataDF$State, str_trim)    # remove surrounding whitespace to state col
```

####4. Create a separate data set with only CO & TX beer and remove IBU's with NA's.
```{r echo = TRUE}
beerCOTX=subset(allBeerDataDF,(allBeerDataDF$State=="CO" | allBeerDataDF$State=="TX") & !is.na(BitternessUnits))
```

####5. Order beerCOTX by IBU (ascending)
```{r echo = TRUE}
beerCOTX <- beerCOTX[with(beerCOTX,order(BitternessUnits)),] # sort by IBU, ascending
```

###B. Create initial plots of the data
####6. Plot Colorado Beers' ABV vs IBU
```{r echo = TRUE}
scatterplot <- ggplot(data=filter(beerCOTX,State=="CO"), aes(AlcoholByVolume,BitternessUnits,col=AlcoholByVolume)) +
  geom_point() + coord_flip() +                                                             # make horizontal scatter plot
  ggtitle("Fig 1 - Colorado ABV and IBU") +                                                 # add title
  theme(plot.title = element_text(color="dark blue",hjust=0.5)) +                           # format main title
  xlab("Alcohol Content") +  ylab("Bitterness") +                                           # add axis labels
  theme(axis.title.x = element_text(color="dark blue")) +                                   # format x title
  theme(axis.title.y = element_text(color="dark blue"))                                     # format y title

#linear fit by lm to scatterplot
scatterplot+geom_smooth(method = "lm", colour = "black", fill="yellow")
```

####6. Plot Texas Beers' ABV vs IBU
```{r echo = TRUE}
scatterplot <- ggplot(data=filter(beerCOTX,State=="TX"), aes(AlcoholByVolume,BitternessUnits,col=AlcoholByVolume)) +
  geom_point() + coord_flip() +                                                             # make horizontal scatter plot
  ggtitle("Fig 2 - Texas ABV and IBU") +                                                    # add title
  theme(plot.title = element_text(color="dark blue",hjust=0.5)) +                           # format main title
  xlab("Alcohol Content") +  ylab("Bitterness") +                                           # add axis labels
  theme(axis.title.x = element_text(color="dark blue")) +                                   # format x title
  theme(axis.title.y = element_text(color="dark blue"))                                     # format y title

#linear fit by lm to scatterplot
scatterplot+geom_smooth(method = "lm", colour = "black", fill="yellow")
```

###C. Model the data
####7. Plot data, add linear model, check assumptions
```{r echo = TRUE}
# creating subset for CO, easier to work with
beerCO=subset(beerCOTX,beerCOTX$State=="CO")

fitCO <- lm(AlcoholByVolume~BitternessUnits,data=beerCO) # linear model
plot(beerCO$AlcoholByVolume~beerCO$BitternessUnits,      # scatter plot for CO
     main="Fig 3 - Colorado ABV vs IBU",                 # add plot title
     xlab="IBU",                                         # add x-axis label
     ylab="ABV")                                         # add y-axis label
abline(fitCO,col='dark green')                           # add linear fit line

residsCO=resid(fitCO)                                    # residuals for CO
plot(beerCO$BitternessUnits,residsCO,                    # plot CO residuals
     main="Fig 4 - Colorado Residual Plot-ABV vs IBU",   # add plot title
     xlab="IBU",                                         # add x-axis label
     ylab="ABV Residuals")                               # add y-axis label
qqnorm(residsCO,main="Fig 5 - Colorado Normal Q-Q Plot", # QQ Plot for CO residuals
       xlab="Normal Score",                              # add y-axis label
       ylab="Residuals")                                 # add x-axis label

# creating subset for TX, easier to work with
beerTX=subset(beerCOTX,beerCOTX$State=="TX") 

fitTX <- lm(AlcoholByVolume~BitternessUnits,data=beerTX) # linear model
plot(beerTX$AlcoholByVolume~beerTX$BitternessUnits,      # scatter plot for TX
     main="Fig 6 - Texas ABV vs IBU",                    # add plot title
     xlab="IBU",                                         # add x-axis label
     ylab="ABV")                                         # add y-axis label
abline(fitCO,col='red')                                  # add linear fit line

residsTX=resid(fitTX)                                    # residuals for TX
plot(beerTX$BitternessUnits,residsTX,                    # plot TX residuals
     main="Fig 7 - Texas Residual Plot-ABV vs IBU",      # add plot title
     xlab="IBU",                                         # add x-axis label
     ylab="ABV Residuals")                               # add y-axis label

qqnorm(residsTX,main="Fig 8 - Texas Normal Q-Q Plot",    # QQ Plot for CO residuals
       xlab="Normal Score",                              # add y-axis label
       ylab="Residuals")                                 # add x-axis label
```

####8. Assumptions for the CO Beer
1. We will assume the data are independent.
2. No patterns seen in the residuals plot so relationship of the data is linear.
3. The variance appears constant in the residuals plot.
4. The normal QQ plot show residuals are relatively normal.

#### Assumptions for the TX Beer
1. We will assume the data are independent.
2. No patterns seen in the residuals plot so relationship of the data is linear.
3. The variance appears constant in the residuals plot.
4. The normal QQ plot show residuals are relatively normal.

###D. Gain inference from the model
####9. Print parameter estimate table. Interpret slope. Is there evidence the relationship between TX & CO ABV vs IBU is significantly different?
#### COLORADO 
```{r echo = TRUE}
summary(fitCO)
```

####For every 1 IBU increase, the estimated ABV increases 3.676e-04 (~0.037%).

#### TEXAS
```{r echo = TRUE}
summary(fitTX)
```

####For every 1 IBU increase, the estimated ABV increases 4.172e-04 (~0.042%).
#### There doesn't appear to be any evidence supporting a significant difference with ABV vs IBU between Colorado and Texas.

####10. Confidence Intervals
#### COLORADO
```{r echo = TRUE}
confint(fitCO,level=.95)
```
####We are 95% confident for every 1 IBU increase, the ABV increase has a plausible range of 2.997e-04 to 4.354e-04 (~0.03% to 0.0435%). 
#### TEXAS
```{r echo = TRUE}
confint(fitTX,level=.95)
```

####We are 95% confident for every 1 IBU increase, the ABV increase has a plausible range of 3.44e-04 to 4.904e-04 (~0.034% to 0.049%). 
####Since there is an overlap with the Colorado and Texas confidence intervals, there is not enought evidence supporting a significant diference with ABV vs IBU betweeen Colorado and Texas.
###E. Compare two competing models: External Cross Validation
####11. Add IBU squared colume to data
```{r echo = TRUE}
beerCOTX <- within(beerCOTX, IBU2 <- beerCOTX$BitternessUnits*beerCOTX$BitternessUnits)
head(beerCOTX)
```

####12. For each state, create a training and test set from the data (60%/40% split respectively). Print a summary of each new data frame. There should be four: TrainingCO, TestCO, TrainingTX, TestTX.
```{r echo = TRUE}
#Divide CO into training and test set into 60% training & 40% test
samplesizeCO=nrow(filter(beerCOTX,State=="CO"))
train_percent = .6
train_indicesCO = sample(seq(1,samplesizeCO,length = samplesizeCO),train_percent*samplesizeCO)
trainCO = subset(beerCOTX,beerCOTX$State=="CO")[train_indicesCO,] # CO random training data from original data
testCO = subset(beerCOTX,beerCOTX$State=="CO")[-train_indicesCO,] # CO random test data from original data
```

####Summary for Colorado Training Set
```{r echo = TRUE}
summary(trainCO)
```

####Summary for Colorado Test Set
```{r echo = TRUE}
summary(testCO)
```

```{r echo = TRUE}
#Divide TX into training and test set into 60% training & 40% test
samplesizeTX=nrow(filter(beerCOTX,State=="TX"))
train_percent = .6
train_indicesTX = sample(seq(1,samplesizeTX,length = samplesizeTX),train_percent*samplesizeTX)
trainTX = subset(beerCOTX,beerCOTX$State=="TX")[train_indicesTX,] # TX random training data from original data
testTX = subset(beerCOTX,beerCOTX$State=="TX")[-train_indicesTX,] # TX random test data from original data
```

####Summary for Texas Training Set
```{r echo = TRUE}
summary(trainTX)
```

####Summary for Texas Test Set
```{r echo = TRUE}
summary(testTX)
```

####13. Compare Linear and Quadratic Models
####Models fit for CO
```{r echo = TRUE}
#Fit the linear model for cO
fitTrainCO1 = lm(AlcoholByVolume~BitternessUnits, data = trainCO) # get the beta hats
predsTrainCO1 = predict(fitTrainCO1)                  # Model predictions on training data which was used to fit the model
predsTestCO1 = predict(fitTrainCO1, newdata = testCO) # Model predictions using test data

#Fit the quadratic model for cO
fitTrainCO2 = lm(AlcoholByVolume~BitternessUnits + IBU2, data = trainCO) # get the beta hats
predsTrainCO2 = predict(fitTrainCO2)                  # Model predictions on training data which was used to fit the model
predsTestCO2 = predict(fitTrainCO2, newdata = testCO) # Model predictions using test data
```

####Calculate the CO MSEs
```{r echo = TRUE}
MSEholderTestCO1 = sum((predsTestCO1 - testCO$AlcoholByVolume)^2)/(length(testCO$AlcoholByVolume) - 2)     # MSE Test, linear
MSEholderTestCO2 = sum((predsTestCO2 - testCO$AlcoholByVolume)^2)/(length(testCO$AlcoholByVolume) - 2)     # MSE Test, quadratic
```
####CO Model Comparison Results
```{r}
paste0("Colorado MSE for the Linear Model Test Data = ",MSEholderTestCO1)
paste0("Colorado MSE for the Quadratic Model Test Data = ",MSEholderTestCO2)
```
####Since the MSE for the linear model is greater than the MSE Delta for the quadratic model, the quadratic model is better than the linear model for Colorado AVB vs IBU. The difference is very small. The next step might be to check if this difference is significant.

####Models fit for TX
```{r echo = TRUE}
#Fit the linear model for TX
fitTrainTX1 = lm(AlcoholByVolume~BitternessUnits, data = trainTX) # get the beta hats
predsTrainTX1 = predict(fitTrainTX1)                  # Model predictions on training data which was used to fit the model
predsTestTX1 = predict(fitTrainTX1, newdata = testTX) # Model predictions using test data

#Fit the quadratic model for cO
fitTrainTX2 = lm(AlcoholByVolume~BitternessUnits + IBU2, data = trainTX) # get the beta hats
predsTrainTX2 = predict(fitTrainTX2)                  # Model predictions on training data which was used to fit the model
predsTestTX2 = predict(fitTrainTX2, newdata = testTX) # Model predictions using test data
```

####Calculate the TX MSEs
```{r echo = TRUE}
MSEholderTestTX1 = sum((predsTestTX1 - testTX$AlcoholByVolume)^2)/(length(testTX$AlcoholByVolume) - 2)     # MSE Test, linear
MSEholderTestTX2 = sum((predsTestTX2 - testTX$AlcoholByVolume)^2)/(length(testTX$AlcoholByVolume) - 2)     # MSE Test, quadratic
```

####TX Model Comparison Results
```{r}
paste0("MSE for the Linear Model Test Data = ",MSEholderTestTX1)
paste0("MSE for the Quadratic Model Test Data = ",MSEholderTestTX2)
```
####Since the MSE for the linear model is greater than the MSE Delta for the quadratic model, the quadratic model is better than the linear model for Texas AVB vs IBU. The difference is very small. The next step might be to check if this difference is significant.

####Github Location: https://github.com/stevebramhall/DoingDataScience/tree/master/HW9
